---
title: "Diversidad y estructura genética de *Pocillopora* spp en Bahía Culebra"
author: "Randall Hidalgo"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    highlight: kate
    fig_width: 11
    fig_height: 8
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, 
                      warning = FALSE, 
                      message = FALSE)
```

Load libraries

```{r}
library(tidyverse)      # v2.0.0
library(adegenet)       # V2.1.10
library(mmod)           # v1.3.3
library(hierfstat)      # v0.5-11
library(poppr)          # v2.9.5
library(pophelper)      # v2.3.1
library(gridExtra)      # v2.3
library(RColorBrewer)   # v1.1-3
library(PopGenReport)   # v3.1
library(patchwork)      # v1.2.0
```

Load Data

```{r}
load(file = "pocillopora_cc.RData")
```


# Genetic Structure

## Global Fst

Observe Nei's Fst (Gst) - by locus and global

```{r}
setPop(pocillopora_cc)<-~Poblacion
bootstrap <- chao_bootstrap(pocillopora_cc)
summarise_bootstrap(bootstrap, statistic = Gst_Nei)
```

## Pairwise Nei's Fst (populations)

Calculate Fst among populations in a pairwise manner

```{r}
# Fst
pares<-pairwise.neifst(pocillopora_cc, diploid = T)
pares

# Confidence interval - bootstrap
boot.ppfst(pocillopora_cc,nboot = 1000)
```


## Discriminant Analysis of Principal Components

This is a graphical way to show clusters of microsatellite data. First, it is necessary to obtain the number of PCs necessary.

```{r}
# Substitue NA values because dudi.pca functions needs no NAs
datos<-tab(pocillopora_cc, freq=T, NA.method="mean")
datos<-tab(pocillopora_cc, freq=T, NA.method="zero")

# Calculate PCA
pca1 = dudi.pca(datos, scannf = FALSE, scale = FALSE, nf = 3)

# Calculate percentage of variance explained by each axis (PCA)
percent = pca1$eig/sum(pca1$eig)*100

# Plot
barplot(percent, ylab = "Variance explicada (%)", ylim = c(0,50),
        names.arg = round(percent, 1))

# crossvalidation
crossval = xvalDapc(datos, pocillopora_cc$pop, 
                    result = "groupMean", xval.plot = TRUE)

# The number of PCA is the one associated with the lowest root mean square error (12 in this case)
crossval$`Root Mean Squared Error by Number of PCs of PCA` %>% 
  round(digits = 3)

numPCs = as.numeric(crossval$`Number of PCs Achieving Lowest MSE`)
numPCs
```

Actual plot

```{r}
dapc1 = dapc(pocillopora_cc, 
             pocillopora_cc$pop, 
             n.pca = numPCs, n.da = 3)

# Create a dataframe with individual coordinates
ind_coords = as.data.frame(dapc1$ind.coord)


# Raname dataframe's columns
colnames(ind_coords) = c("Axis1","Axis2")

# Add a columns with individuals
ind_coords$Ind = indNames(pocillopora_cc)
# Add column with populations
ind_coords$Site = pocillopora_cc$pop

# Calculate centroid position (mean) for each population
centroid = aggregate(cbind(Axis1, Axis2) ~ Site, 
                     data = ind_coords, FUN = mean)

# Add centroid's coordinates to ind_coords df
ind_coords = left_join(ind_coords, centroid, 
                       by = "Site", suffix = c("",".cen"))

# Define colors
cols = brewer.pal(nPop(pocillopora_cc), "Set2")

# Customize labels
xlab = paste("Axis 1 (", format(round(percent[1], 1), 
                                nsmall=1)," %)", sep="")
ylab = paste("Axis 2 (", format(round(percent[2], 1), 
                                nsmall=1)," %)", sep="")



# Scatterplot using axis1 vs. axis2
ggplot(data = ind_coords, aes(x = Axis1, y = Axis2))+
  geom_hline(yintercept = 0)+
  geom_vline(xintercept = 0)+
  # spider segments
  geom_segment(aes(xend = Axis1.cen, yend = Axis2.cen, 
                   colour = Site), show.legend = F)+
  # points
  geom_point(aes(fill = Site), shape = 21, 
             size = 3, show.legend = F)+
  # Centroids
  geom_label(data = centroid, aes(label = Site, fill = Site), 
             size = 4, show.legend = F)+
  # Colors
  scale_fill_manual(values = cols)+
  scale_colour_manual(values = cols)+
  # Axis
  labs(x = xlab, y = ylab)+
  ggtitle("DAPC")+
  theme_bw()+
  theme(axis.text = element_text(size = 15))
```

```{r, eval=FALSE, echo=FALSE}
tiff(filename = "dapc.tiff", width = 1500*2, height = 1000*2, res = 300)
dev.off()
```


## STRUCTURE Software

### Evanno's Method

```{r}
# STRUCTURE files (do not use this command to read local files)
sfiles <- 
  list.files(path="structure_pd3_clone", 
             full.names=T) # <-- Look carefully here! You must put in the path where you have put the results files. No other files can be in that directory!
# basic usage
slist <- readQ(files=sfiles, filetype="structure")

sr1 <- summariseQ(tabulateQ(slist))

p <- evannoMethodStructure(data=sr1,exportplot=F,returnplot=T,returndata=F,basesize=12,linesize=0.7)
grid.arrange(p)
```

### STRUCTURE plot

```{r}
# plot

cols = c("#66C2A5", "#FC8D62", "#8DA0CB")

# Load data K = 2
struc_k2<-read.csv("Structure_plots/clone_structure_R_2.csv", sep = ";")
names(struc_k2)<-c("ID", "missing", "pop", 
                "cluster1", "cluster2")

# Load data K = 3
struc_k3<-read.csv("Structure_plots/clone_structure_R_3.csv", sep = ",")
names(struc_k3)<-c("ID", "missing", "pop", 
                "cluster1", "cluster2", "cluster3")

struc2<-pivot_longer(data = struc_k2, 
             cols = c(cluster1, cluster2),
             names_to = "cluster", 
             values_to = "probabilidad")

struc3<-pivot_longer(data = struc_k3, 
             cols = c(cluster1, cluster2, cluster3),
             names_to = "cluster", 
             values_to = "probabilidad")

struc2<-struc2 %>% 
  mutate(ID=as.factor(ID), probabilidad=as.numeric(probabilidad),
         cluster=as.factor(cluster), pop=as.factor(pop))

struc3<-struc3 %>% 
  mutate(ID=as.factor(ID), probabilidad=as.numeric(probabilidad),
         cluster=as.factor(cluster), pop=as.factor(pop))


K2 <- struc2 %>% 
  ggplot(aes(ID, probabilidad, fill=cluster))+
  geom_col(width = 1)+
  ylab("Probabilidad")+
  geom_vline(xintercept = c(3.5, 15.5))+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = cols[2:1])+
  theme(axis.text.x = element_blank(),
        legend.position = "none", 
        #axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

K3 <- struc3 %>% 
  ggplot(aes(ID, probabilidad, fill=cluster))+
  geom_col(width = 1)+
  ylab("Probabilidad")+
  geom_vline(xintercept = c(3.5, 15.5))+
  scale_y_continuous(expand = c(0,0))+
  scale_fill_manual(values = cols[c(1,2,3)])+
  scale_x_discrete(breaks = c("J15", "M15", "Z15"),
                   labels = c("Jícaro", "Matapalo", "Zapotillal"))+
  theme(axis.text.x = element_text(size = 20),
        legend.position = "none", 
        #axis.text.x = element_text(angle = 90),
        axis.title.x = element_blank(),
        axis.ticks.x = element_blank())

plots <- K2 / K3 + plot_layout(axes = "collect")
plots
```


